Characterizing subarctic European vegetation's response on key environmental variables for land surface modeling of ozone pollution and risk assessment

### 2021-10-04 VegOzone.tex
Refine text

In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions at the NIBIO operated Environment Centre Svanhovd (Sør-Varanger, NOR). The NIBIO centre Svanhovd focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e. $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths. We investigate the 2018/19 growing season in a climatological context. To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$. Svanhovd synonymously refers to the atmospheric monitoring site Svanvik/Pasvik throughout the paper.\\

@@

The location of the NIBIO environment centre Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate. The 1992--2012 climatologies of temperature and precipitation are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}). Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band. We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation. Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars. Months with average temperatures below freezing are denoted with a star. As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}. The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February. The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred in July 16 the same year.

@@

\section{DO3SE modeling}
\label{sec:do3se}
In the following, we investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk. We model ozone uptake of natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model \citep{ACP:Bueker2012} and develop bespoke parameterizations for common plant species in our focus area. First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results. The relevant equations of the $\mathrm{DO_3SE}$ model are summarized in Appendix~\ref{appendix:do3se_model}.

@@

We deduced the dominant species on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees). Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following. We base our parameterization of coniferous trees on the MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for central Europe. For a comprehensive list of model default parameters, consult Supplement~Table~S1.

Initial test simulations with the MM PFTs revealed an unrealistically low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}). Substantial stomatal conductance occurred only during an extended warm period in late July. In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality. We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance in perennial grassland. Therefore, we saw the necessity to propose bespoke parameterizations describing an acclimation of the default PFTs to the climatic conditions in the focus area. Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments.

We assume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season. Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (see Appendix~\ref{appendix:do3se_model} for details) to acclimate the growth potential to the local climate. Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these.

We define two different hypothetical temperature acclimations: subarctic and cold. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM. This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to default values. In the same way, \emph{subarctic} constructed to represent a species which is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures. Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000. In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature PDF of the GS. The most extreme of the two acclimation types is \emph{subarctic}.
In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{temp}$ for perennial grassland together with temperature PDFs deduced from five months GS in two climatological periods. The periods 1990s and 2000s have been chosen to indicate that the vegetation is subject to ongoing climate change Thus, the predominant acclimation at present is not known.

@@

Regarding the opening extent of stomata at low light conditions, we look at the PPFD threshold for a $50\,\unit{\%}$ opening of plants' stomata. We assume that this threshold may differ in subarctic species compared to species in less extreme climates. We derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) analytically for each PFT or species $k$
\begin{equation}
  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
  \label{eq:inverse_function}
\end{equation}
First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ at $0.5$ 
\begin{equation}
  \gamma^k(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}^k},
  \label{eq:inverse_function_halfway}
\end{equation}
and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$ and invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ to find $\alpha'(\gamma')$. The resulting functions for perennial grassland (PPFD0.8, PPFD1.2) are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b). All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

@@

As indicated above, a better acclimation to prevailing climate conditions is characterized by a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. To compare our bespoke parameterizations, we propose the following metric. We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of meteorological data and apply mean and standard deviation. We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth). In additional, a small deviation between noon and morning and a low standard deviation would indicate a higher robustness to variability in growing conditions. Based on the criteria, we identify \emph{subarctic}--PPFD0.8 as the best parameterization for all PFTs (Fig.~\ref{fig:javis_func_opt_mean}). As expected due to the small adjustments, coniferous trees display the smallest differences between the different types of acclimation, while the differences for perennial grassland are substantial as a response to the stronger proposed temperature acclimation.

@@

Another important factor for ozone exposure metrics is the timing and length of the GS. Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$. The net photosynthesis vs. time of year data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019. $A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively. This value will be used for all PFTs alike. The resulting growing season for coniferous trees in 2019 was one month longer than in 2018. For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
For comparison, the default values for central Europe ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season. Due to a lack of quantitative field observation at Svanhovd, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland. This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.

\begin{table}[t]
  \caption{Start and end of growing season in units of \unit{doy}.}
  \label{tab:sensitivity_tests_gs}
  \begin{tabular}{lcll}
    \tophline
    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
    \middlehline
    Default$^\star$ & -- & 100 & 307\\
    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^\dagger$ \\
    & 2019 & 130$^*$ & 274$^\dagger$ \\
    \multirow{2}{*}{Coniferous tree} & 2018 & 122$^\dagger$ & 261$^\dagger$ \\
    & 2019 & 106$^\dagger$ & 274$^\dagger$ \\
    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^\ddagger$ & 261$^\dagger$\\
    & 2019 & 161$^\ddagger$ & 274$^\dagger$ \\
    \bottomhline
  \end{tabular}
  \belowtable{$^\dagger$ MODIS (Aqua/Terra) net photosynthesis product;\\
    $^\star$ \citet{ACP:Simpson2012},\\
    $^*$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
    $^\ddagger$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
\end{table}

Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}. The soil at Svanhovd is characterized in an ICP Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP Forest soil classification system (V. Timmerman, personal communication, 2021 May 10) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.

The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions. Leaves of downy birch in northern Fennoscandia are probably smaller than the $5\,\unit{cm}$ from the MM parameterization of silver birch leaves. A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$. In addition, \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area. Both factors will affect leaf boundary layer resistance, the smaller leaves would tend to reduce leaf boundary layer resistances whilst the shorter trees would reduce canopy roughness and increase canopy boundary layer resistances and hence the uptake of \chem{O_3}. In addition, one Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping. Heights were measured in 2004, when the stand was 90 years old. Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V. Timmerman, personal communication, 2021 May 10).

@@

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) response functions. Canopy height and leaf size are adjusted to local values in the model simulations with bespoke growing season (GS besp.). The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of systematic uncertainties arising from the use of generic parameterizations established from less climatically extreme conditions.
For comparison, critical levels for reduction in biomass are shown as horizontal dashed lines \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021}. We identify that an increase in cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction. The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018. Due to the shape of the light response function, a variation of the extent of stomatal opening at low light intensities leads to an asymmetric response in $\mathrm{POD_1}$ and can cause a higher sensitivity to drought conditions. Similarly, the temperature acclimation to cold or subarctic conditions can lead to an amplification of effects by droughts.

The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here. Our bespoke temperature parameterizations suggest considerable underestimations of ozone risk for subarctic species when relying on the generic MM parameterizations.

For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Deciduous trees display the largest variation of ozone uptake among all PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light parameterization is pronounced. With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses. However, taking the bespoke GS into account, this is reversed for the \emph{subarctic} acclimation type. This can be explained by more favorable growing conditions in 2019. Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke GS. The temperature acclimation-dependent difference in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Coniferous trees' parameterization was already well acclimated to prevailing temperatures from the start and therefore differs least. Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018 for both bespoke temperature response functions. The overall difference deduced from bespoke GS for subarctic and cold types with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

>>>>

In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions at the NIBIO operated Environment Centre Svanhovd (Sør-Varanger, NOR). The NIBIO Environment Centre Svanhovd focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e. $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths. We investigate the 2018/19 growing season in a climatological context. To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$. Svanhovd synonymously refers to the atmospheric monitoring site Svanvik/Pasvik throughout the paper.\\

@@

The location of the NIBIO Environment Centre Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate. The 1992--2012 climatologies of temperature and precipitation are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}). Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band. We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation. Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars. Months with average temperatures below freezing are denoted with a star. As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}. The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February. The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred in July 16 the same year.

@@

\section{DO3SE modeling}
\label{sec:do3se}
We investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk. We model ozone uptake of natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model \citep{ACP:Bueker2012} and develop bespoke parameterizations for common plant species in our focus area. First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results. The relevant components of the $\mathrm{DO_3SE}$ model are summarized in Appendix~\ref{appendix:do3se_model}.

@@

We deduced the dominant species on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees). Generic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following. Our parameterization of coniferous trees is based on the MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for central Europe. For a comprehensive list of generic MM model parameters, consult Supplement~Table~S1.

Initial test simulations with the MM PFTs revealed an unrealistically low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}). Substantial stomatal conductance in this species occurred only during an extended warm period in late July. In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality. We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance in perennial grassland. Therefore, we saw the necessity to propose bespoke parameterizations describing an acclimation of the generic PFTs to the climatic conditions in the focus area. Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments.

We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season. Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ to acclimate the growth potential to the local climate. For details regarding these functions see Appendix~\ref{appendix:do3se_model}. Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these.

We define two different hypothetical temperature acclimations: subarctic and cold. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM. This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to MM values. In the same way, \emph{subarctic} is constructed to represent a species which is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures. Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000. In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature PDF during the GS. The most extreme of the two acclimation types is \emph{subarctic}.
In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{temp}$ for perennial grassland together with temperature PDFs deduced from the five months GS in two climatological periods. The periods 1990s and 2000s have been chosen to indicate that the vegetation has been subject to climate change. Thus, the predominant acclimation at present is not known.

@@

Regarding the acclimation of $f_\mathrm{light}$, we presume that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates. We, therefore, look at the PPFD threshold for a $50\,\unit{\%}$ opening of plants' stomata. We derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) analytically for each PFT or species $k$
\begin{equation}
  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
  \label{eq:inverse_function}
\end{equation}
First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ at $0.5$ 
\begin{equation}
  \gamma^k(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}^k},
  \label{eq:inverse_function_halfway}
\end{equation}
and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. By inverting Eq.~(\ref{eq:inverse_function}) for $\alpha$, we find $\alpha'(\gamma')$. The resulting functions for perennial grassland (PPFD0.8, PPFD1.2) are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b. All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

@@

As indicated above, a better acclimation to prevailing climate conditions is characterized by a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. To compare our bespoke parameterizations, we propose the following metric. We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of key environmental factors' data and apply mean and standard deviation. We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth). In addition, a small deviation between noon and morning and a low standard deviation would indicate a higher robustness to variability in growing conditions. Based on these criteria, we identify \emph{subarctic}--PPFD0.8 as the best parameterization for all PFTs (Fig.~\ref{fig:javis_func_opt_mean}). As expected due to the small adjustments, coniferous trees display the smallest differences between the different parameterizations, while the differences for perennial grassland are substantial as a response to the proposed temperature acclimation.

@@

Another important factor is the timing and length of the GS. Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product \citep{MODIS_PSN} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$. The net photosynthesis vs. time of year data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019. $A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively. This value will be used for all PFTs alike. The resulting growing season for coniferous trees in 2019 is, hence, one month longer than in 2018. For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR). For comparison, the generic values for central Europe ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season. Due to a lack of quantitative field observation at Svanhovd, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland. This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.

\begin{table}[t]
  \caption{Start and end of growing season in units of \unit{doy}.}
  \label{tab:sensitivity_tests_gs}
  \begin{tabular}{lcll}
    \tophline
    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
    \middlehline
    Default$^b$ & -- & 100 & 307\\
    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^a$ \\
    & 2019 & 130$^c$ & 274$^a$ \\
    \multirow{2}{*}{Coniferous tree} & 2018 & 122$^a$ & 261$^a$ \\
    & 2019 & 106$^a$ & 274$^a$ \\
    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^d$ & 261$^a$\\
    & 2019 & 161$^d$ & 274$^a$ \\
    \bottomhline
  \end{tabular}
  \belowtable{$^a$ MODIS (Aqua/Terra) net photosynthesis product;\\
    $^b$ \citet{ACP:Simpson2012},\\
    $^c$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
    $^d$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
\end{table}

Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}. The soil at Svanhovd is characterized in an ICP Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP Forest soil classification system (V. Timmerman, personal communication, May 2021) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.

The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions. A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$ which is smaller than in the MM parameterization. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area. One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping. Heights were measured in 2004, when the stand was 90 years old. Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V. Timmerman, personal communication, 2021 May 10).

@@

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations. Canopy height and leaf size are adjusted to local values in the model simulations with bespoke growing season (GS besp.). The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of generic parameterizations established from less climatically extreme conditions.
Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. We identify that an increase in cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction. The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018. Due to the shape of $f_\mathrm{light}$, a symmetric variation leads to an asymmetric response in $\mathrm{POD_1}$. We find that an opening of stomata at lower light intensities can cause a higher sensitivity to drought conditions. Similarly, a temperature acclimation to cold or subarctic conditions can lead to an amplification of effects by droughts.

The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here. Our bespoke temperature parameterizations suggest considerable underestimations of ozone risk for subarctic species when relying on the generic MM parameterizations.

For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Deciduous trees display the largest variation of ozone uptake among all PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light parameterization is pronounced. With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses. However, taking the bespoke GS into account, this is reversed for the subarctic temperature acclimation. This can be explained by more favorable growing conditions in 2019. Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke GS. The temperature acclimation-dependent difference in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). The MM Boreal Norway spruce arameterization was already well acclimated to prevailing temperatures and therefore differs least from our bespoke parameterizations. Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018. The overall difference deduced from bespoke GS for subarctic and cold parameterizations with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
